---
title: "Atorus-Test-Drive"
author: "Eli Miller"
date: "10/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Tplyr)
library(dplyr)
library(huxtable)
library(magrittr)
library(pharmaRTF)
library(knitr)
load("data/adae.Rdata")
load("data/adlb.Rdata")
load("data/adsl.Rdata")
```

## Tplyr

Here is a Table displaying the end of study flag and the disposition code across the study.

A lot of cool stuff is happening here. Be sure to check out the Tplyr Booth!

```{r cars}
eos_tab <- tplyr_table(adsl, TRT01A) %>%
  add_total_group() %>%
  add_layer(
    group_count(EOSSTT) %>%
      add_total_row()
  ) %>%
  add_layer(
    group_count(vars(EOSSTT, DCDECOD)) %>%
      set_nest_count(TRUE)
  )
eos_b <- eos_tab %>%
  build() %>%
  apply_row_masks(row_breaks = TRUE) %>%
  select(-starts_with("ord")) %>%
  select(c(1,2,5,4,3)) %>%
  add_column_headers(
    "  | Placebo \\line(N=**Placebo**) | Xanomeline Low Dose \\line(N=**Xanomeline Low Dose**) | Xanomeline High Dose \\line(N=**Xanomeline High Dose**) | Total \\line(N=**Total**)",
    header_n(eos_tab)
  )

kable(eos_b)
```

## pharmaRTF

Huxtable does the heavy lifting. Here were passing some styling information to the table. We pass the table back to huxtable to make the rtf for the table.

```{r huxtable_prep}
  # The Column names are in the table already so don't add them.
hx <- huxtable::as_hux(eos_b, add_colnames = FALSE) %>%
  # We use some encoding in the column headers so don't escape that
  huxtable::set_escape_contents(FALSE) %>%
  # Make the top row bold
  huxtable::set_bold(1, 1:ncol(eos_b), TRUE) %>%
  # Center top row
  huxtable::set_align(1, 1:ncol(eos_b), "center") %>%
  # Set the width of the entire table
  huxtable::set_width(1.5)

```

Here pharmaRTF is adding titles and footnotes to the huxtable table. We note several additional styling arguments for the titles and column headers.

```{r pressure}
rtf <- pharmaRTF::rtf_doc(hx) %>%
  # Add a couple of titles
  pharmaRTF::add_titles(
    #  first title, centered, size 30 font
    pharmaRTF::hf_line("The Atorus Test Drive",
                       align = "center", bold = TRUE, font_size = 30),
    # second title, size 14 font split alignment
    pharmaRTF::hf_line(c("An End of Study Status Table", "Created with Tplyr, huxtable, and pharmaRTF"),
                       align = "split", italic = TRUE, font_size = 14)) %>%
  pharmaRTF::add_footnotes(
    # footnote to contain the page number and the date the rtf was made
    pharmaRTF::hf_line(c("PAGE_FORMAT: Page %s or %s", "DATE_FORMAT: %H:%M %A, %B %d, %Y"),
                       align = "split")
    ) %>%
  # Make the table appear more compact
  pharmaRTF::set_ignore_cell_padding(TRUE) %>%
  # Set a buffer of one row above and below the column headers
  pharmaRTF::set_column_header_buffer(top = 1, bottom = 1) %>%
  # Set the default font size as 10
  pharmaRTF::set_font_size(10) %>%
  # Output RTF.
  suppressWarnings(write_rtf("test1.rtf"))
```

## Tplyr

Again Tplyr is used to create the table itself. Additional factors are added to the adsl. These will appear in the output even though they aren't actually in the data!

```{r}
adsl <- adsl %>% 
  mutate(
    SEX = recode(SEX, M = "Male", F = "Female"), 
    RACE = factor(RACE, c("AMERICAN INDIAN OR ALASKA NATIVE", "ASIAN", "BLACK OR AFRICAN AMERICAN", 
                          "NATIVE HAWAIIN OR OTHER PACIFIC ISLANDER", "WHITE", "MULTIPLE"))
  )

t <- tplyr_table(adsl, TRT01P) %>% 
  add_total_group() %>% 
  add_layer(name = 'Sex', 
    group_count(SEX, by = "Sex n (%)") %>% 
      set_missing_count(f_str('xx', n), string = "NA") %>%
      set_denom_ignore("NA")
  ) %>% 
  add_layer(name = 'Age',
    group_desc(AGE, by = "Age (Years)")
  ) %>% 
  add_layer(name = 'Age group', 
    group_count(AGEGR1, by = "Age Categories n (%)") %>% 
      set_missing_count(f_str('xx', n), string = "NA") %>%
      set_denom_ignore("NA")
  ) %>% 
  add_layer(name = 'Race', 
    group_count(RACE, by = "Race n (%)") %>% 
      set_missing_count(f_str('xx', n), string = "NA") %>%
      set_denom_ignore("NA") %>% 
      set_order_count_method("byfactor")
  ) %>% 
  add_layer(name = 'Ethnic', 
    group_count(ETHNIC, by = "Ethnicity n (%)") %>% 
      set_missing_count(f_str('xx', n), string = "NA") %>%
      set_denom_ignore("NA")
  )

dat <- build(t) %>%
  apply_row_masks(row_breaks = TRUE) %>%
  select(-starts_with("ord")) %>%
  select(1,2,3,6,5,4) %>%
  add_column_headers(
    " | | Placebo\\line(N=**Placebo**) | Xanomeline Low Dose\\line(N=**Xanomeline Low Dose**) | Xanomeline High Dose\\line(N=**Xanomeline High Dose**) | Total\\line(N=**Total**)",
    header_n=header_n(t)
  )

kable(dat)
```

## pharmaRTF

```{r}
  # The Column names are in the table already so don't add them.
hx <- huxtable::as_hux(dat, add_colnames = FALSE) %>%
  # We use some encoding in the column headers so don't escape that
  huxtable::set_escape_contents(FALSE) %>%
  # Make the top row bold
  huxtable::set_bold(1, 1:ncol(dat), TRUE) %>%
  huxtable::set_align(1, 1:ncol(dat), "center") %>%
  huxtable::set_width(1.5)
```

We've added a couple extra things to this document. It's longer and appears across multiple pages; In the bottom left the footnote will display which page you're on. We've also added a footnote to display the file path the document was created by, improving tractability.

```{r}
rtf <- rtf_doc(hx)  %>%
  pharmaRTF::add_titles(
    pharmaRTF::hf_line("The Atorus Test Drive",
                       align = "center", bold = TRUE, font_size = 30),
    pharmaRTF::hf_line(c("An End of Study Status Table", "Created with Tplyr, huxtable, and pharmaRTF"),
                       align = "split", italic = TRUE, font_size = 14),
    replace = TRUE
    ) %>%
  pharmaRTF::add_footnotes(
    pharmaRTF::hf_line(c("PAGE_FORMAT: Page %s or %s", "DATE_FORMAT: %H:%M %A, %B %d, %Y"),
                       align = "split"),
    pharmaRTF::hf_line("FILE_PATH: Sourced from: %s"),
    replace = TRUE
    ) %>%
  set_ignore_cell_padding(TRUE) %>%
  set_column_header_buffer(top = 1, bottom = 1) %>%
  set_font_size(10) %>%
  set_pagesize(c(height = 6, width = 12)) %>%
  write_rtf("test2.rtf")
```
