---
title: "Tplyr Test Drive"
author: "Mike Stackhouse"
date: "10/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(dplyr)
require(magrittr)
require(Tplyr)
require(kableExtra)
load('data/adsl.Rdata')
load('data/adae.Rdata')
load('data/adlb.Rdata')
```

## Take Tplyr for a Test Drive

Welcome! We've created this repository and document to allow you to get some hands on experience using our packages [Tplyr](https://atorus-research.github.io/Tplyr/index.html) and [pharmaRTF](https://atorus-research.github.io/pharmaRTF/index.html). If you're viewing this on GitHub, then we recommend you hop on over to RStudio Cloud through [this link](https://rstudio.cloud/project/1744479) to test drive the package for yourself without the hassle of setting things up for yourself. 

Here we have a few hands on examples of Tplyr for you to use, explore, and experiment. For more in depth explanations of certain aspects of this package, check out our package website right [here](https://atorus-research.github.io/Tplyr/index.html). 

So without further ado...

## Example 1: Demographics

Let's start with a basic demographics table, and let's make it no frills. This will start introducing you to the concept of stacking layers. For simplicity, let's also start by looking at sex, age, age category, and race.

```{r Demographics 1}
# Create the table object
t <- tplyr_table(adsl, TRT01P, where = SAFFL == "Y") %>% 
  # Add a count layer for sex
  add_layer(
    group_count(SEX, by = "Sex n (%)")
  ) %>% 
  # Add a descriptive statistics layer for age
  add_layer(
    group_desc(AGE, by = "Age (years)")
  ) %>% 
  # Add a count layer for age categories
  add_layer(
    group_count(AGEGR1, by = "Age Categories n (%)")
  ) %>% 
  # Add a count layer for race
  add_layer(
    group_count(RACE, by = "Race n (%)") 
  )

t %>% 
  build() %>% 
  # Kable and all forward is just for presentation in the HTML
  kable() %>% 
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")
```

## Example 2: Demographics Cleaned Up

That looks pretty good, but there's a few things that we can clean up a bit more:

- Let's spell out the sexes explicitly
- The order for the age categories is off
- Let's assume that we want to show a few more races for which we had no subjects

Factors can help us out here - let's see how. This can be handled easily with a bit of pre-processing.

```{r Demographics 2}
# Do some pre-processing
adsl <- adsl %>% 
  mutate(
    # Expand out sex to full words
    SEX = recode(SEX, M = "Male", F = "Female"), 
    # Convert RACE to a factor and add levels to be displayed
    RACE = factor(RACE, c("AMERICAN INDIAN OR ALASKA NATIVE", "ASIAN", "BLACK OR AFRICAN AMERICAN", 
                          "NATIVE HAWAIIN OR OTHER PACIFIC ISLANDER", "WHITE", "MULTIPLE")),
    # Convert age group to a factor and order properly
    AGEGR1 = factor(AGEGR1, c('<65', '65-80', '>80'))
  )

# Create the table object and do everything like example 1
t <- tplyr_table(adsl, TRT01P, where = SAFFL == "Y") %>% 
  add_layer(
    group_count(SEX, by = "Sex n (%)")
  ) %>% 
  add_layer(
    group_desc(AGE, by = "Age (years)")
  ) %>% 
  add_layer(
    group_count(AGEGR1, by = "Age Categories n (%)")
  ) %>% 
  add_layer(
    group_count(RACE, by = "Race n (%)") 
  )

t %>% 
  build() %>% 
  kable() %>% 
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")

```

## Example 3: Beyond the Defaults

Let's change things up a bit. 

- Let's use sex as a column variable, so we see the subjects within sex by treatment group. 
- Let's also switch up the summary statistics performed on age, and take out the percentages in all of the count summaries.
- Let's also add a total group into the mix

```{r Demographics 3}
# Create the table object, this time with an additional columns argument
t <- tplyr_table(adsl, TRT01P, where = SAFFL == "Y", cols=SEX) %>% 
  # Add a total group that includes all treated subjects
  add_total_group() %>% 
  # Set default count formats for all layers
  set_count_layer_formats(
    n_counts = f_str("xx", n)
  ) %>% 
  # Add a descriptive statistics layer for age
  add_layer(
    group_desc(AGE, by = "Age (years)") %>% 
      # Set the format strings and use auto precision
      set_format_strings(
        "Mean (SD)" = f_str("a.a (a.a+1)", mean, sd),
        "Q1, Median, Q3" = f_str("a.a+1, a.a+1, a.a+1", q1, median, q3)
      )
  ) %>% 
  # And the other layers like before, but standard count layer formats
  # pass through from the table level
  add_layer(
    group_count(AGEGR1, by = "Age Categories n (%)")
  ) %>% 
  add_layer(
    group_count(RACE, by = "Race n (%)") 
  )

t %>% 
  build() %>% 
  kable() %>% 
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")
```

## Example 4: Adverse Events

Let's consider another type of table - Adverse Events. 

Here we demonstrate a few other capabilities of count layers specifically:

- Two variables can be counted together using the `vars()` argument. Both will be summarized, with the second variable nested inside the first.
- Distinct counts can be calculated with a specified grouping variable
- Distinct and non-distinct counts can be presented side by side
- Order variables can controlled with helper functions, and for count layers, users are able to specify particular number from a selected column from within the summary to extract out to an order variable for the layer
- Risk difference can be added by specifying treatment groups to compare. Currently this uses the `prop.test()` function from base R, but we plan to expand this in the future to add common methods.


```{r Adverse Events 1}
# Create the table object
t <- tplyr_table(adae, TRTA, where = SAFFL == "Y") %>% 
  # Set a separate population data set
  set_pop_data(adsl) %>% 
  # Set the population dat aset treatment variable
  set_pop_treat_var(TRT01A) %>% 
  # Add a count layer
  add_layer(
    # Add a nested count layer by supplying two variables to be summarized
    group_count(vars(AEBODSYS, AEDECOD)) %>% 
      # Set distinct counts by subject
      set_distinct_by(USUBJID) %>% 
      # Set the formats to include distinct counts, distinct percent, and event counts
      set_format_strings(
        n_counts = f_str("xx (xx.x%) [x]", distinct, distinct_pct, n)
        ) %>% 
      # Get rid of the outer row label column
      set_nest_count(TRUE) %>% 
      # Order using counts from the summarized data
      set_order_count_method('bycount') %>% 
      # Use the distinct_n variable as that order variable
      set_result_order_var(distinct_n) %>% 
      # Pull the data from the Xanomeline High Dose column
      set_ordering_cols('Xanomeline High Dose') %>% 
      # Add risk difference calculations for both treatment groups vs. placebo
      add_risk_diff(
        c('Xanomeline High Dose', 'Placebo'), 
        c('Xanomeline Low Dose', 'Placebo')
      )
  )

# The warnings here come from calculating risk difference 
# when there's really not enough data to be statistically significant,
# so we'll just mask them right now
suppressWarnings(build(t)) %>% 
  # For presentation sake let's order the columns
  select(starts_with('row'), starts_with('var'), starts_with('rdiff'), starts_with('ord')) %>% 
  kable() %>% 
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")
```

## Example 5: Shift Tables

Let's do one last example - shift tables! This leverages the last layer type that we haven't touched on yet. So what do you need to make a shift table? Variable to mark a change in state. 

There's another prep step here that might be necessary. If you don't have a representation of each state in your data for both variables, then that row or column will not form. You can get around this by using factor variables and specifying the levels. This also assists with ordering.

Note that here we're also using `by` variables, of which you can submit multiple by using `vars()`. What we end up getting here is the summary of changes is state by parameter and visit. And for simplicity's sake, we've filtered down to just the Creatine Kinase test.

```{r Shift Table 1}
# Convert the 'from' and 'to' columns to factors, order and force missing levels
adlb$ANRIND <- factor(adlb$ANRIND, levels=c("L", "N", "H"))
adlb$BNRIND <- factor(adlb$BNRIND, levels=c("L", "N", "H"))

# Create the table object
t <- tplyr_table(adlb, TRTA, where=PARAMCD == "CK") %>%
  # Add the shift layer, which takes two variables that will be the 
  # row and column variable you want for presentation
  # Additionally note here that we're using two by variables to group by
  # parameter and visit (though we've filtered to the CK parameter)
  add_layer(
    group_shift(vars(row = BNRIND, column = ANRIND), by = vars(PARAM, VISIT))
  )

t %>% 
  build() %>% 
  kable() %>% 
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")
```